---
title: "Seroprevalence analysis"
author: '2020-08-21'
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
  fig.path = "figures/",
  out.width = "100%"
)
```

```{r}
cat('Date-time: ', Sys.time())
## Load libraries
library(ggplot2)
library(lubridate)
library(dplyr)
library(readr)
library(gsheet)
library(anytime)
library(seroprevalence)
library(databrew) # devtools::install_github('databrew/databrew')
# options(scipen = '999')
ggplot2::theme_set(theme_bw())
library(yaml)
creds <- yaml::yaml.load_file('credentials.yaml')
library(haven)
library(sp)
```

```{r}
# Read in data
if('fetched_data.RData' %in% dir()){
  load('fetched_data.RData')
} else {
  data <- odk_get_data(url = creds$url,
                       id = 'seroprevalence',
                       unknown_id2 = TRUE,
                       user = creds$user,
                       password = creds$password)
  # Get locations
  pd <- data$non_repeats
  # extract lat long
  pd$lon <- as.numeric(unlist(lapply(strsplit(pd$gps, ' '), function(x) x[2])))
  pd$lat <- as.numeric(unlist(lapply(strsplit(pd$gps, ' '), function(x) x[1])))
  
  # Read in OSM data
  dat <- sf::st_read('data/map.osm', layer = 'multipolygons', quiet = TRUE)
  dat_sp <- as(dat, 'Spatial')
  dat_sp <- dat_sp[is.na(dat_sp@data$type) | dat_sp@data$type == 'multipolygon',]
  
  dat2 <- sf::st_read('data/map2.osm', layer = 'multipolygons', quiet = TRUE)
  dat_sp2 <- as(dat2, 'Spatial')
  dat_sp2 <- dat_sp2[is.na(dat_sp2@data$type) | dat_sp2@data$type == 'multipolygon',]
  
  dat3 <- sf::st_read('data/map3.osm', layer = 'multipolygons', quiet = TRUE)
  dat_sp3 <- as(dat3, 'Spatial')
  dat_sp3 <- dat_sp3[is.na(dat_sp3@data$type) | dat_sp3@data$type == 'multipolygon',]
  
  dat_sp <- rbind(dat_sp3, dat_sp2, dat_sp)
  dat_sp <- dat_sp[!duplicated(dat_sp@data),]
  
  # Read in Marta's DTA
  df <- haven::read_dta('Cizur_all_working_data_20200813.dta') %>%
    mutate(infection = ifelse(overall_infection == 1, TRUE, FALSE))
  # Remove extreme observations
  df <- df %>%
    filter(!is.na(lon)) %>%
    filter(lon >= -1.9, lon <= 1.5,
           lat >= 42.7, lat <= 42.9)
  
  rounder <- 3
  # Make a basic grid map
  xr <- range(df$lon, na.rm = T)
  yr <- range(df$lat, na.rm = T)
  xr <- round(xr, rounder)
  yr <- round(yr, rounder)
  xs <- seq(min(xr), max(xr), by = as.numeric(paste0('.', paste0(rep(0, abs(rounder)-1), collapse = ''), 1)))
  ys <- seq(min(yr), max(yr), by = as.numeric(paste0('.', paste0(rep(0, abs(rounder)-1), collapse = ''), 1)))
  griddy <- expand.grid(x = xs,
                        y = ys) %>%
    mutate(positives = NA,
           total = NA,
           smoothed = NA)
  # Get a rounded version of locations
  df <- df %>%
    mutate(lon_round = round(lon, rounder),
           lat_round = round(lat, rounder))
  # Get spatial version
  df_sp <- df
  coordinates(df_sp) <- ~lon+lat
  proj4string(df_sp) <- proj4string(bohemia::moz0)
  griddy_sp <- griddy
  coordinates(griddy_sp) <- ~x+y
  proj4string(griddy_sp) <- proj4string(df_sp)
  distance_matrix <- rgeos::gDistance(spgeom1 = df_sp, 
                                      spgeom2 = griddy_sp,
                                      byid = TRUE)
  # Prepare data
  for(i in 1:nrow(griddy)){
    sub_data <- df %>%
      filter(lon_round == griddy$x[i],
             lat_round == griddy$y[i])
    if(nrow(sub_data)>0){
      griddy$positives[i] <- length(which(sub_data$infection))
      griddy$total[i] <- nrow(sub_data)
    }
      this_distances <- as.numeric(distance_matrix[i,])
      vals <- weighted.mean(as.numeric(df$infection), w  = 1 / this_distances^(1/2))
      griddy$smoothed[i] <- vals
  }

  dat_sp@data$id <- 1:nrow(dat_sp@data)
  dat_fort <- fortify(dat_sp, region = 'id')
  
  
  save(data,
       dat,
       dat_fort,
       dat_sp,
       griddy,
       griddy_sp,
       df,
       df_sp,
       pd,
       data,
       file = 'fetched_data.RData')
}
```


*IMPORTANT*: `r length(which(is.na(pd$lon)))` location values were not collected of a total of `r nrow(pd)` participants (ie, `r round(length(which(is.na(pd$lon))) / nrow(pd) * 100, digits = 1)` %).

```{r}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = df_sp[!df_sp@data$infection,], radius = 1) %>%
  addCircleMarkers(data = df_sp[df_sp@data$infection,], color = 'red', radius = 2)
```

```{r, fig.height = 13, fig.width = 13}
plot(dat_sp)
points(df_sp[!df_sp@data$infection,], col = 'blue', pch = 16)
points(df_sp[df_sp@data$infection,], col = 'red', pch = 16)
```

```{r}
library(raster)
r <- raster::rasterFromXYZ(xyz = griddy[,c('x', 'y', 'smoothed')])
proj4string(r) <- proj4string(bohemia::moz0)
center <- apply(coordinates(r), 2, median)
bb <- bbox(r)
pal <- colorNumeric(c('blue', 'lightblue',
                                                              # 'beige',
                                                              # 'grey',
                                                              # 'yellow',
                                                              # 'white',
                                                              # 'orange',
                                                              'darkorange', 
                                                              'red',
                                                              'darkred',
                                                              'black'), values(r),
  na.color = "transparent")
leaflet() %>% addProviderTiles('Stamen.Toner') %>%
    addRasterImage(r, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal, values = values(r),
    title = "Smoothed estimated prevalence") %>%
  leaflet::setMaxBounds(lng1 = bb[1,1], lat1 = bb[2,1],
                        lng2 = bb[1,2], lat2 = bb[2,2])

```

```{r}
ggplot(data = griddy,
       aes(x = x,
           y = y,
           fill = smoothed)) +
  geom_tile() +
  geom_polygon(data = dat_fort,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               lwd = 0.2,
               color = 'black') +
  databrew::theme_simple() +
  scale_fill_gradientn(name = '', colors = colorRampPalette(c('lightblue',
                                                              # 'beige',
                                                              # 'grey',
                                                              # 'yellow',
                                                              # 'white',
                                                              # 'orange',
                                                              'darkorange', 
                                                              'red',
                                                              'darkred'))(20)) +
  xlim(-1.82, -1.62) +
  ylim(42.75, 42.81) +
  ggthemes::theme_map()
```

# Clustering

Kulldorf clustering detection method

Kulldorff spatial cluster detection method for a study region with n areas. The method constructs zones by consecutively aggregating nearest-neighboring areas until a proportion of the total study population is included. Given the observed number of cases, the likelihood of each zone is computed using either binomial or poisson likelihoods. The procedure reports the zone that is the most likely cluster and generates significance measures via Monte Carlo sampling. Further, secondary clusters, whose Monte Carlo p-values are below the α-threshold, are reported as well.

```{r}
library(class)
library(geosphere) # has functionality for spatial clustering
df_spatial <- df_sp
df_spatial@data <- data.frame(df_spatial@data)
mdist <- distm(df_spatial)
test <- hclust(as.dist(mdist), method = "complete")

# Define a distance threshold
d = 1000 # meters (change this up or down to make bigger or smaller clusters, respectively)
df_spatial$cluster <- cutree(test, h = d)
# df_spatial$cluster <- if_else(df_spatial$latitude>-25.42, 1, 2)


# View our clustering results
library(RColorBrewer)
cols <- colorRampPalette(brewer.pal(9, "Spectral"))(length(unique(df_spatial$cluster)))
cols <- sample(cols, length(cols))
df_spatial$color <- cols[df_spatial$cluster]
plot(df_spatial, col = df_spatial$color, pch = 1, axes = TRUE, cex.axis = 0.5, las = 1)

# How many clusters
length(unique(df_spatial$cluster))

# Now that we are satisfied with our clustering results,
# we will create polygons with borders
clusters <- unique(sort(df_spatial$cluster))

# Loop through each cluster, and for each one use "convex hull"
# to create a border
ch <- list()
for (i in 1:length(clusters)){
  this_cluster <- clusters[i]
  sub_df <- df_spatial[df_spatial@data$cluster == this_cluster,]
  x <- rgeos::gConvexHull(sub_df)
  ch[[i]] <- x
}


# Create delaunay triangulation / voronoi tiles for entire surface
voronoi <- function(shp = df_spatial){
  
  shp@data <- data.frame(shp@data)
  
  # Fix row names
  row.names(shp) <- 1:nrow(shp)
  
  # Remove any identical ones
  shp <- shp[!duplicated(shp$lon,
                         shp$lat),]
  
  # Helper function to create coronoi polygons (tesselation, not delaunay triangles)
  # http://carsonfarmer.com/2009/09/voronoi-polygons-with-r/
  voronoipolygons = function(layer) {
    require(deldir)
    require(rgeos)
    crds = layer@coords
    z = deldir(crds[,1], crds[,2])
    w = tile.list(z)
    polys = vector(mode='list', length=length(w))
    require(sp)
    for (i in seq(along=polys)) {
      pcrds = cbind(w[[i]]$x, w[[i]]$y)
      pcrds = rbind(pcrds, pcrds[1,])
      polys[[i]] = Polygons(list(Polygon(pcrds)), ID=as.character(i))
    }
    SP = SpatialPolygons(polys)
    voronoi = SpatialPolygonsDataFrame(SP, data=data.frame(x=crds[,1], 
                                                           y=crds[,2], row.names=sapply(slot(SP, 'polygons'), 
                                                                                        function(x) slot(x, 'ID'))))
  }
  # http://gis.stackexchange.com/questions/180682/merge-a-list-of-spatial-polygon-objects-in-r
  appendSpatialPolygons <- function(x) {
    ## loop over list of polygons
    for (i in 2:length(x)) {
      # create initial output polygon
      if (i == 2) {
        out <- maptools::spRbind(x[[i-1]], x[[i]])
        # append all following polygons to output polygon  
      } else {
        out <- maptools::spRbind(out, x[[i]])
      }
    }
    return(out)
  }
  
  tile_polys <- voronoipolygons(shp)
  # Add the bairro numbers
  tile_polys@data$cluster_number <- the_clusters <- shp$cluster
  cols <- rainbow(as.numeric(factor(tile_polys@data$cluster_number)))
  
  # Disolve borders
  x = gUnaryUnion(tile_polys, id = tile_polys$cluster_number)
  
  jdata = SpatialPolygonsDataFrame(Sr=x, 
                                   data=data.frame(cluster_number = as.numeric(as.character(names(x)))),FALSE)
  
  return(jdata)
}

# Get voronoi tesselations
dfv <- voronoi(shp = df_spatial)

# For each cluster in df_spatial, we want to get the % positive
get_positivity <- function(var = "infection"){
  require(dplyr)
  x <- df_spatial@data
  x$var <- df_spatial@data[,var]
  out <- x %>%
    group_by(cluster) %>%
    summarise(positives = length(which(var == 1)),
              total = n()) %>%
    mutate(percentage = positives / total * 100)
  return(out)
}

# Run our function
get_positivity("infection")

# Define a function for identifying hotspots
ana_hotspot <- function(var = "infection",
                        plot_it = FALSE,
                        seed = 1){
  # set.seed(seed)  
  # Use the kulldorf method to get hotspots
  library(SpatialEpi)
  
  # Create centroids
  centroids <- as.matrix(coordinates(dfv))
  
  # Get the positivity rate for a variable
  positivity <- get_positivity(var)
  
  # Define some parameters about our results
  cases <- positivity$positives
  population <- positivity$total
  n_strata <- nrow(dfv@data)
  
  expected_cases <- sum(cases, na.rm = TRUE)
  expected_cases <- expected_cases * (population / sum(population, na.rm = TRUE))
  
  # Set paramaters
  pop.upper.bound <- 0.5
  n.simulations <- 9999
  alpha.level <- 0.05
  plot <- FALSE
  
  poisson <- kulldorff(geo = centroids,
                       cases = cases,
                       population = population,
                       expected.cases = expected_cases,
                       pop.upper.bound = pop.upper.bound,
                       n.simulations = n.simulations,
                       alpha.level = 0.05,
                       plot = FALSE)
  
  if(plot_it){
    # get clusters
    cluster <- poisson$most.likely.cluster$location.IDs.included
    # cluster <- df_spatial@data[cluster,"cluster"]
    secondary_cluster <- poisson$secondary.clusters$location.IDs.included
    
    # Plot
    plot(dfv,axes=TRUE, border = adjustcolor("black", alpha.f = 0.5))
    plot(dfv[cluster,],add=TRUE,col="red")
    # plot(man3, add = TRUE)
    points(df_spatial, pch= ifelse(df_spatial@data[,var] == 1, "+", "0"),
           cex=0.5,
           col= adjustcolor("black", alpha.f = 0.5))
    if(!is.null(secondary_cluster)){
      plot(dfv[secondary_cluster,], add = TRUE, col = "yellow")
    }
    p_value <- poisson$most.likely.cluster$p.value
    p_value <- ifelse(p_value == 1, "> 0.99", p_value)
    title(main = paste0(var, " p:",p_value ))
  } else {
    poisson
  }
}

# Run the function just to make a pretty map of hotspots
ana_hotspot("infection", plot_it = TRUE)
```


